---
title: "Your Turn : R essentials"
author: "Your netid, last name, first name here"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

# 1. Console vs .Rmd
Compared to working in the R console, .Rmd allows you to rerun (reproduce) code much more effortlessly


`r bfcolor("Exercise 1:", "blue")` **R as a calculator:** use any mathematical operators (+, -, /, and *) to create an expression and make sure it works as expected. _What is `2+2`, `100*3`, `100/10`, `sqrt(25)`?_

`r bfcolor("Solutions:", "blue")`


`r bfcolor("Exercise 2:", "blue")` **Object Assignment:** Assign the value of your mathematical operation to the variable `y`. Note the change in the environment pane. 

`r bfcolor("Solutions:", "blue")`

`r bfcolor("Exercise 3:", "blue")` **Load** the `palmerpenguins` **package** so that we have access to all functions and data in this package. 

`r bfcolor("Solutions:", "blue")`


`r bfcolor("Exercise 4:", "blue")` **View** the `penguins` **dataset**

`r bfcolor("Solutions:", "blue")`

`r bfcolor("Exercise 5:", "blue")` Take a look at the `flipper_length` variable


`r bfcolor("Solutions:", "blue")`

`r bfcolor("Exercise 6:", "blue")` **Find the average** `flipper_length`

`r bfcolor("Solutions:", "blue")`


`r bfcolor("Exercise 7:", "blue")` **Get help** on the mean **function**

`r bfcolor("Solutions:", "blue")`


# 2.  Packages


**OYO:**

1. Check if the package `datasets` is installed, loaded?

2. Check if the package `nycflights13` is loaded. If not loaded, load it.



## 3. R essentials : Defining objects

```{r}
ysq <- 2^2 # note the change in the environment

ysq
```
```{r error = TRUE}
y <- 3
# use previously defined object/variable y 
#and assign it's cube to the object xsq
y_cube <- y^3 

y_cube
```
### R is case sensitive.

```{r error = TRUE}
YSQ
```


### Pay attention to naming conventions


## 4. R essentials : More flexibility in printing objects

```{r echo=TRUE, eval = TRUE, error = TRUE}
# without print function:
#R prints values using default options
ysq
print(ysq)
```


```{r echo=TRUE, eval = TRUE}
# Printing quoted text
welcome_msg <- "Welcome to PSTAT10"
welcome_msg
# Printing using print function
print(welcome_msg2 <- "Welcome to PSTAT10, Fall 22 with Dr. Ravat")
```
```{r echo=TRUE, eval = TRUE}
# quote = FALSE for no quotes
print(welcome_msg3 <- "Welcome to PSTAT10", quote = FALSE)
```

`r bfcolor("OYO: " ,"orange")`
Check out help page for `?print` for more options and examples


## 5. Environments




