---
title: "PSTAT 10: Worksheet 13"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Due 30 minutes after Lab Section, on Canvas"
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
---


\newcommand{\Prob}{\mathbb{P}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

---

# Instructions for Submission

- We are back to having only computing exercises on the worksheets. As such, this worksheet is due **30 minutes after the end of Lab**, as was the case in the first few weeks of this course.

\

---

# Exercises

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 


\

---

For this worksheet, we need will again connect to the Chinook DB from lecture. Make sure the database is available to this .Rmd file in `data` folder

## Exercise 1: Establishing a Connection

Load the necessary packages and open a connection to the Chinook database housed in the sqlite file in the data folder

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

---

## Exercise 2: Terminology
You may use Google to find explanations that make sense to you. 

**(a)** Briefly explain the purpose of primary keys and foreign keys in a relational database.

`r bfcolor("Solutions:", "blue")`


**Replace this line with your answer**


\

**(b)** Briefly explain how relational databases model real world data.

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer**


\

**(c)** Briefly explain the difference between SQL, SQLite, and `RSQLite`.

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer**


\

---


## Exercise  3: Tables and Foreign keys

**(a)** List all tables in the database

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer**


\

**(b)** **List all of the foreign key relationships in the database.** How many are there in all?

Write out these relationships in the following format:

`TableName.ForeignKey -> TargetTableName.PrimaryKey`.

For example, the following indicates that
`SupportRepId` in the `Customer` table is a foreign key to the primary key
of the `Employee` table, which is `EmployeeId.`

1. `Customer.SupportRepId -> Employee.EmployeeId`

*Hint:*
There are 11 total relationships to list.
Do not simply "guess" by looking at the diagram from lecture. Access
the Chinook database schema information using the SQLite `pragma` command.


`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer**


\


---


## Exercise 4: SQL queries

**(a)** Select the `CustomerId`, `FirstName`, `LastName`, `State`, `Country` of all customers living in California.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(b)** How many customers are from USA? 

_Hint_: Use `count(*)` in your select query

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

---

## Exercise 5: SQL queries

Choose another table from the database. Write a SQL query and describe in words what your query is doing. 

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

---

# Last exercise : Learning Gains

Please remember to fill out these learning gains before submitting your work to Canvas!

-   Gain 1
-   Gain 2
-   Gain 3
