---
title: "PSTAT 10: Worksheet 5"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Due on Canvas 30 mins after Section following lecture."
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
params:
  solution: TRUE
  release: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```
---

# Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 

---

# Exercise 1: 


Take a look at the `starwars` dataset in three  ways. For plots in this exercise, do something other than what we did in lecture. 

**(a)** Use `glimpse` and  `str()` function and look at the help file. What difference do you notice between information given by these three functions. Make sure to load any packages you will need


`r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```



**(b)** Use `tail`, `head` with appropriate arguments to look at the first three and last three rows respectively. Use the help for these functions if needed

`r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```

**(c)** What are all the column names?

`r bfcolor("Solutions:", "blue")`


```{r}
# Your Code here
```



**(d)** How much data do we have? Find it in two ways. 

`r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```


**(e)** Pick a single numerical variable and make a histogram of it.

`r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```

**(f)** Pick two numerical variables from the data set and make a scatter plot. You may need to remove `NA` values. Describe the relation, if any, between these two variables. 

`r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```



**(g)** Pick a single categorical variable from the data set and make a bar plot of its distribution. You may need to omit `NA` values.

`r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```

**(h)** Pick a numerical variable and a categorical variable and make a visualization (you pick the type!) to visualize the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

 `r bfcolor("Solutions:", "blue")`

```{r}
# Your Code here
```

---

# Last exercise : Reflecting

Write at least one thing you struggled with but figured out 

- Gain 1
- Gain 2
- Gain 3

Write at least one thing you still have a question about

- question 1
- question 2
- question 3

Write at least one way in which you will resolve this question

- resolution 1
- resolution 2
- resolution 3


