---
title: "PSTAT 10: Worksheet 2"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Due on Canvas 30 mins after Section following Lecture 2"
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```
---

## Exercise 0: YAML


Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 

---

## Exercise 1: Element-ary, Dear Watson

Create a vector `x` containing the elements 5, 7, 3, 10, 1, 7, 19 ,5, 2

Write code to:

**(a)** extract the 4th element of `x`.

`r bfcolor("Solutions:", "blue")`



**(b)** extract the 5th , 6th and 7th elements of `x`

`r bfcolor("Solutions:", "blue")`



**(c)** find the median of `x`

`r bfcolor("Solutions:", "blue")`



**(d)** determine the length of `x`

`r bfcolor("Solutions:", "blue")`



**(e)** **(Optional)** Assign the name of nine of your friends as names attribute of this vector `x`. Verify that that the names have been assigned correctly.

`r bfcolor("Solutions:", "blue")`


\

---

## Exercise 2: Rep-and-Roll

Using `rep()` function, create the following vectors:

**(a)** 6 6 6 6 6 6 6

`r bfcolor("Solutions:", "blue")`



**(b)** 3 3 3 4 4 4 6 6 6

`r bfcolor("Solutions:", "blue")`



\

---

## Exercise 3: Combined Intellect

Given variables `p <- c(2,3,4,6,6)` and `q <- c(7,2,4,3,2)`. 

**(a)** Create a vector `z` by combining the vectors `p` and `q`.

`r bfcolor("Solutions:", "blue")`



**(b)** Explain the difference in the results of `p + q` and
`sum(p + q)`

`r bfcolor("Solutions:", "blue")`



\

---

## Exercise 4: :)


Rewrite the following code using the colon operator.

```{r, eval = F}
x <- c(-50, -51, -52, -53, -54, -53, -52, -51, -50)
```

`r bfcolor("Solutions:", "blue")`



\

---

## Exercise 5: Rep-and-Roll, Electric Boogaloo


Use the replicate function `rep()` to generate the following sequence:

1 1 2 2 3

(Note: there is not an error here. The required sequence has only one instance of ‘3’)

`r bfcolor("Solutions:", "blue")`



---


# Last Exercise: Learning Gains

Please remember to fill out these learning gains before submitting your work to Canvas!

- Gain 1
- Gain 2
- Gain 3
