---
title: "PSTAT 10: Worksheet 8"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Follow the instructions on Canvas for submitting!"
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

---

# Instructions for Submission

- This worksheet consists of two parts: Computing Exercises, and Theory Exercises. 
- You must complete the computing exercises by filling in your code on this `.Rmd` file, just like you did with the previous worksheets. 
- You must complete the theory exercises on a separate sheet of paper.
- The submission portal on Canvas will give you instructions on where and how to submit these two documents (your knitted `.html` file, along with a scan or photo of your handwritten work.)
- Because the submission for this worksheet is a bit more involved than previous ones, you will have until **11:59pm on the day of your Section** to upload your work.

\

---

# Computing Exercises

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 


\

---

## Exercise 1: Seeds

- The **`set.seed()`** function allows us to reproduce the exact sample each time we execute any function involving randomness. The syntax is `set.seed(n)`, where `n` is a seed number.

The seed number you choose is the starting point used in the generation of a sequence of random numbers, which is why you’ll obtain the same results given the same seed number.


**(a)** Create a sample of 10 flips of a fair coin. Use zero and one to denote heads and tails respectively.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(b)** Run the above code chunk three times. Do you observe the same or  a different set of 10 flips in each repetition? 

`r bfcolor("Solutions:", "blue")`

**Replace this line with a description indicating you've carried out the necessary steps.**


\

**(c)** In the first line of the chunk below, set the seed to your favorite number. Then, copy-paste the code from part (a) below your call to `set.seed()`. Run this code chunk three times. Do you observe the same or different set of 10 flips each time. Briefly explain.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

**Replace this line with a description of your results.**

\

---

## Exercise 2: Probability Distributions

The probability distribution of a discrete random variable $X$ is as shown in the table:

| x | 1 | 2 | 3 | 4 | 5 |
|---|:-:|:-:|:-:|:-:|:-:|
| $\mathbb{P}(X = x)$ | 0.0 | $a$ | 0.4 | 0.1 | 0.03 | 

Use `R` for these calculations

**(a)** Find the missing value $a$.

`r bfcolor("Solution:", "blue")`

```{r}
## Add your code here
```


\

**(b)** Compute $\mathbb{P}(X < 4)$. (You should plug in the value for $a$ you found in part (a); that is, your final answer should NOT be in terms of $a$.)

`r bfcolor("Solution:", "blue")`

```{r}
## Add your code here
```

\

**(c)** Find the mean of the random variable $X$. (You should plug in the value for $a$ you found in part (a); that is, your final answer should NOT be in terms of $a$.)

`r bfcolor("Solution:", "blue")`

```{r}
## Add your code here
```

\

---

**(d)** Construct the cumulative distribution function of $X$. (You should plug in the value for $a$ you found in part (a); that is, your final answer should NOT be in terms of $a$.)

`r bfcolor("Solution:", "blue")`

```{r}
## Add your code here
```

\

---

# Theory Exercise:

## Exercise 3: Coin Flips

Consider flipping two coins and counting number of heads.  **Do these calculations by hand; do not use `R` for these calculations!**

(a) Write down the p.m.f. 
(b) Compute the expected value $\mathbb{E}[X]$, variance $\mathrm{Var}(X)$, and cumulative distribution function of $X$.

\

---

# Last exercise : Learning Gains

Please remember to fill out these learning gains before submitting your work to Canvas!

-   Gain 1
-   Gain 2
-   Gain 3

