---
title: "Your Turn : Data structures in `R`"
author: "Your netid, last name, first name here"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

# 1. Matrices

## Examples: Create and subset a matrix

```{r}
(x <- seq(-5, 5, 2))
x_mat <- matrix(x, nrow=2, ncol=3) #by columns
x_mat
```

```{r}
dim(x_mat)
```
##
```{r}
colnames(x_mat) <- c("Col 1", "Col 2", "Col 3")  
rownames(x_mat) <- c("Row 1", "Row 2")
x_mat
```
```{r}
x_mat[ , c("Col 1","Col 2")]
```

```{r }
dimnames(x_mat)
```


## Creating a matrix by binding to rows or columns
```{r}
(r_mat <- rbind(c(-5, -1, 3), c(-3, 1, 5)))
(c_mat <- cbind(c(-5, -3), c(-1, 1), c(3, 5)))
```

## Matrix subsetting

- Using [ ] provide an  index to each dimension. 
    - order in a matrix is given as [ row , col]

```{r}
(x <- matrix(1:4, nrow=2, ncol=2))
x[1, 1]
x[2, 1]
```
##
**Omitting** an index returns all elements in that dimension
```{r}
x[ , 2]
x[1, ]
x[,-1]
```

##  YT01

```{r}
(x <- matrix(1:9, nrow=3, ncol=3))
```

##  YT02

Create a $4\times 5$ matrix `mx` of integers 1 through 17 (inclusive).
Print the matrix. 

1. Subset `mx` to create the following matrix:
```{r}
#write your code here
```

2. Set all entries of `mx` greater than 10 to zero.

```{r}
#write your code here
```




 



# 2. Arrays

## 1. Create an array

```{r}
A = array(data = 1:18, dim = c(3, 3, 2))
```

```{r}
A
dim(A)
```

## 2. Modify the array to include: 
   a) columns names; 
   b) row names;
   c) layer names
```{r}
dimnames(A) = list(c("R1","R2","R3"),
                   c("C1","C2","C3"),
                   c("L1","L2"))
```

```{r}
A
```
```{r}
dimnames(A)
```

## 3. On your own: practice subsetting using [ ] and dim names 

```{r}
A[, "C1", "L1"]
```


## 4. `apply` function


### 4a.  row sums for A
```{r}
apply(A, 1, sum)
rowSums(A)
```

### 4b.  col sums for A
```{r}
# write your code here
```

### 4c. Sum of all  cells

```{r}
apply(A, c(1, 2), sum)
```


# 3. Factors

## Example 1: Eye color
```{r}
eyes <- factor(x = c("blue", "green", "green"))
eyes
```


```{r}
eyes <- factor(x = c("blue", "green", "green"),
               levels = c("blue", "brown", "green"))
eyes
```

### What are the levels?
```{r}
levels(eyes) # extracting levels
```

### How many levels are there?
```{r}
nlevels(eyes)
```
### Produce a frequency distribution of the data.
```{r}
table(eyes)
```

### Produce a plot of the data.
```{r}
plot(eyes)
```

## Ordered factor :Two options in `R`

### 1. `ordered()` function for ordinal data.
```{r}
eyes
class(eyes)
eyes <- ordered(eyes)
eyes
class(eyes)
```


### 2. `factor()` function with the argument `ordered = TRUE`
```{r}
 # Create ordered factor
my_fac_ordered <- factor(c("c", "a", "b", "c", "a"), 
                          levels = c("c", "a", "b"),
                          ordered = TRUE)
my_fac_ordered
class(my_fac_ordered)
```






# 4. Logical

```{r}
?state
# library(datasets) 
```

1. How many US states have an area less than 10,000 **OR** greater than 100,000 square miles?

```{r}
which( state.area < 10000 | state.area > 10000 )
length(which( state.area < 10000 | state.area > 10000 ))
```

2. How many US states have an area less than 10,000 **AND** greater than 100,000 square miles?

```{r}
which( state.area < 10000 & state.area > 10000)
length(which( state.area < 10000 & state.area > 10000))
```

