---
title: "PSTAT 10: Worksheet 3"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Due on Canvas 30 mins after Section following lecture."
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
params:
  solution: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

# Exercise 1: Matrices

**(a)** Generate a 5 x 4 numeric matrix with elements 1 through 20

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```




**(b)** Select the 3rd column of the matrix

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```




**(c)** Select rows 2,3,4 of columns 1,2,3 of the matrix

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(d)** Add row names and column names to the matrix. 

```{r}
# Write your code here
```



**(e)** Extract the elements in column one by it's column name.

```{r}
# Write your code here
```

---

# Exercise 2: More matrix 

**(a)** Create vectors: a = (1,2,4,5,6); b = (3,2,4,1,9); d = (7,5,5,6,4)

`r bfcolor("Solutions:", "blue")` 

```{r}
# Write your code here
```



**(b)** Without using matrix function, write one line of code to generate a matrix which has rows given by vectors a, b, c in that order. 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```


**(c)** What is the dimension of the matrix you created in (b) above?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```

---

# Exercise 3: Array

```{r}
aaa =  c(2,3,4,6)
bbb = c(5,6,10,12,45)
```

**(a)** Using the above vectors, create an array of two 3 x 4 matrices.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(b)** Sum the rows 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```




**(c)** Sum the columns

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```




**(d)** Sum the cells

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```

---

# Exercise 4: Factor

**(a)** Create a factor with elements: male, female, female, male 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```


**(b)** What are the levels?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(c)** How many levels are there?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(d)** Determine if  this is an ordered factor? Explain your reasoning after your code. 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```


# (Optional) Exercise : Review of Single Logical operator


**(a)** Test whether 2 is less than 4

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(b)** Is 5 equal to 6?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(c)** Is 5+6<4?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```

---

# Exercise 5: Logicals 

**(a)** In which package can we find datasets with information about US states?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```


**(b)** How many US states have an area less than 10 000 square miles? 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```



**(c)** How many US states have an area greater than 100 000 square miles?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here
```




---

# Last exercise : Learning Gains

Please remember to fill out these learning gains before submitting your work to Canvas!

-   Gain 1
-   Gain 2
-   Gain 3

