---
title: 'PSTAT 10: Homework 7'
subtitle: 'YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID'
date: "Due on Canvas by 11:59pm on Week 10 Thursday December 1"
output:
  html_document:
    theme: readable
    toc: yes
---

\newcommand{\Prob}{\mathbb{P}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
library(kableExtra)
```

---

# Instructions for Submission

- As usual, 5 Multiple Choice Questions also appear on Canvas and are a required part of this homework. **Please note** that the multiple choice questions for this homework cover material from throughout the course; not just material tested on this homework!
- Remember to save your `.Rmd` file as `hw07-yournetid.Rmd`

\

---

# Introduction

## Fall 2022 Course Offerings Database

We will be exploring the `F22_Offerings` database, which contains information on a portion of the PSTAT courses being offerred this quarter (Fall 2022). A complete database description is included in a file called `Database_Description.pdf`.


\

---

# Exercises

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 


\

---


## Exercise 1: Creating the Database

You will notice that we have provided the database to you in the form of several `.csv` files. Our first task will be to combine these into a single `.sqlite` database. This will very closely mirror the process we used in Lecture 16. \

**(a)** Load the required libraries for connecting to an `RSQLite` database.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(b)** You should see a file called `f22_offerings.sqlite` included in the `data` folder of this homework. (It should be blank at first!) Establish a connection to the database and assign this connection to a variable called `f22_offerings_db`.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(c)** In the `data` folder, you should see a subfolder called `f22_offerings`. In this subfolder, you should see a series of `.csv` files. Create appropriately-named data frames from these `.csv` files. (Again: Lecture 16 will be very useful!)

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(d)** Write the data frames you created in part (c) to the `f22_offerings.sqlite` database.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(e)** Check that your tables have been successfully written to the database.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

---

## Exercise 2: Exploring the Database

**(a)** Write code to ensure foreign keys constraints are enforced. (This was done several times in Lecture.)

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(b)** List the fields in the `OFFERINGS` relation, and check that they match the fields listed in the database description. 

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(c)** How many courses is Professor Wainwright teaching? **Answer using CODE;** don't just manually count!

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(d)** How many lecture sections of PSTAT 5A are there? **Answer using CODE;** don't just manually count!

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(e)** How many classes have a lecture that meet on Mondays (not necessarily *just* on Mondays)? **Answer using CODE;** don't just manually count!

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(f)** Which professors have offices in South Hall? Display the professors' last and first names, along with their office number. **Answer using CODE!** 

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

---

## Exercise 3: Exploring the Database Further

**(a)** How many TA's are hosting a Section of PSTAT 5A?

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(b)** How many TA's are hosting a Section of PSTAT 10?

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(c)** Which TA's have a first initial between E and I, inclusive? Display the associated TA's Last name as well as their first initial and the class they are TA'ing. **Hint:** Look up how to use the `BETWEEN` clause in `SQL`.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(d)** For each course being offered, identify the TA with the longest last name. Display the course, last name, and first initial of the associated TA's. **Hint:** Look up how to use the `LENGTH` clause in `SQL`.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(e)** Which classes have Calculus as a prerequisite?

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(f)** Create a barplot to visualize the distribution of prerequisites. That is, your x-axis should include the names of all prerequisite subjects (e.g. Calculus, Linear Algebra, etc.) and your y-axis should count the number of courses that require each subject as a prerequisite. **Hint:** Consider extracting all of the prerequisites, using the `strsplit()` and `unlist()` functions, and then calling `barplot()` as we did before.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

---

## Exercise 4: Aggregations and Explorations


**(a)** Produce the following table using a suitable join. 

| CourseID | Days | Times | Lecture_Location | Prof_Surname | Prof_Forename | Office |
|-------:|-------:|-------:|-------:|-------:|-------:|-------:|
| PSTAT 5A | TR | 8-915 | CHEMISTRY 1179 | Swenson | Julie | SH 5524 |
| PSTAT 5A | MWF | 2-250 | EMBARCADERO HALL | Wainwright | Brian | UNKNOWN |
| PSTAT 5H | M | 11-1150 | GIRVETZ 2127 | Meiring | Wendy | SH 5510 |
| PSTAT 8 | TR | 11-1215 | HSSB 1173 | Solis | Sharon | ETR 103B |
| PSTAT 10 | TR | 8-915 | IV THEATRE 1 | Ravat | Uma | SH 5503 |
| PSTAT 115 | TR | 8-915 | HSSB 1173 | Targino | Rodrigo | OG 1203 |
| PSTAT 120A | MWF | 12-1250 | LOTTE LEHMANN CONCERT HALL | Wainwright | Brian | UNKNOWN |
| PSTAT 120A | TR | 2-315 | TD-W 1701 | Coburn | Katie | SH 5524 |

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(b)** Are any professors sharing an office? **Do NOT simply look up the answer from the table you just created in part (a).** Rather, think about how you can answer this question using code. As a hint: think about how you can use `COUNT` and uniqueness to approach this problem. Once you're done, you can check your answer (visually) using the table you created in part (a). **Note:** you will need to examine the result of two separate queries to answer this problem.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

---

## Exercise 5: Fixing a Mistake

You may notice that the `OFFERINGS` relation includes a tuple on PSTAT 105, taught by Dr. Carter, whereas the `PROFESSORS` relation is missing Dr. Carter's information. Let's fix that!

**(a)** Add the following tuple to the `PROFESSORS` relation:

- \<Carter, Andrew, SH 5507\>

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(b)** Check that your tuple has been successfully added.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

**(c)** Now, recount the number of professors that have an office in South Hall.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

---

## Exercise 6: Probabilities

For each of these parts, use a combination of `SQL` queries and computations done in `R`.

**(a)** Suppose a Teaching Assistant is selected at random. Given that they are a TA for PSTAT 120A, what is the probability that their first name begins with the letter "S"?

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\ 

**(b)** In this problem, we will work toward answering the question:

> From the pool of professors included in the dataset, a single professor is to be selected at random. What is the probability that this professor is either teaching PSTAT 5A or has an office in South Hall? 


We will work toward our answer using a combination of the probability theory and `SQL` knowledge we've learnt in this class. 

\

**(i)** Define events, and identify the probability rules you'll use to calculate the desired probability. 

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer.**

\

**(ii)** Notice that the professors' offices and the courses they are teaching are located in two separate relations. Join these two relations, and return all rows; save this resulting table in a data framed called `table1_join_table2`. called `joined_tab`.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```


\

**(iii)** Now, write your `table1_join_table2` data frame from part (ii) above to a new table in the `F22_Offerings` database, called `JOINED_TABLE`.

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\


**(iv)** Finally, use your newly written table to obtain the desired probability by writing as many `SQL` queries are necessary. Keep in mind you may need to perform some computations in `R` (i.e. outside of `SQL` queries).

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

---

## Exericse 7: Closing the Connection

Don't forget to close your connection to the database!

`r bfcolor("Solutions:", "blue")`

```{r}
## Add your code here
```

\

---